Data Preparation

### 2019-01-10 Data Preparation.Rmd

#### Tidy Text Format
We first create two tibbles `d_overview_tidy` and `d_description_tidy` that respectively store the overviews and the descriptions in the tidy text format, with one row per course-year-word.
```{r tidy text data}
d_overview_tidy <- d_description %>%
  select(Code, `Calendar Year`, Overview) %>%
  unnest_tokens(output = word, input = Overview)
print(d_overview_tidy)

@@

#### Stemming
We then use the function `hunspell__stem`, which returns valid stems for a given word, to stem the words in the tibbles `d_overview_tidy` and `d_description_tidy`. We first create a function `stem_hunspell` which, given a term, returns its most basic stem (the last one from the list of stem returned by `hunspell__stem`). We then apply `stem_hunspell` on the words of `d_overview_tidy`. Since, `stem_hunspell` is not a vectorized function and the number of words in `d_overview_tidy` is large, we first use distinct on the `word` variable to find the a list containing *once* each term present in the course overviews and apply the function `stem_hunspell` on this list. We save the terms and their stems in the tibble `stems`. Since the course descriptions can be conceptualized as a subset of the course overviews, the tibble `stems` contains all the terms present in the descriptions and can therefore safely used to stem the words in `d_description_tidy`.
```{r function stem_hunspell, cache = TRUE}
stem_hunspell <- function(term) {
  # look up the term in the dictionary
  stems <- hunspell_stem(term)[[1]]

  if (length(stems) == 0) { # if there are no stems, use the original term
    stem <- term
  } else { # if there are multiple stems, use the last one
    stem <- stems[[length(stems)]]
  }

  return(stem)
}

stems <- d_overview_tidy %>% 
  select(word) %>%
  distinct %>%
  arrange(word) %>%
  mutate(word_stem = purrr::map_chr(.x = word,
                                    .f = stem_hunspell))

# terms for which the stem differs from the original word
stems %>%
  filter(word != word_stem)
```

```{r stemming}
d_overview_stem <- d_overview_tidy %>%
  left_join(stems, by = "word") %>%
  rename(word_original = word,
         word = word_stem) %>%
  filter(!word %in% stop_words$word)
print(d_overview_stem) # See humanities - humanity

d_description_stem <- d_description_tidy %>%
  left_join(stems, by = "word") %>%
  rename(word_original = word,
         word = word_stem) %>%
  filter(!word %in% stop_words$word)
print(d_description_stem) # See starting - start
```

>>>>

#### Tidy Text Format
We first create two tibbles `d_overview_tidy` and `d_description_tidy` that respectively store the overviews and the descriptions in the tidy text format, with one row per course-year-word.
```{r tidy d_description}
d_overview_tidy <- d_description %>%
  select(Code, `Calendar Year`, Overview) %>%
  unnest_tokens(output = word, input = Overview)
print(d_overview_tidy)

@@

#### Stemming
We then use the function `hunspell__stem`, which returns valid stems for a given word, to stem the words in the tibbles `d_overview_tidy` and `d_description_tidy`. We first create a function `stem_hunspell` which, given a term, returns its most basic stem (the last one from the list of stem returned by `hunspell__stem`). We then apply `stem_hunspell` on the words of `d_overview_tidy`. Since, `stem_hunspell` is not a vectorized function and the number of words in `d_overview_tidy` is large, we first use distinct on the `word` variable to find the a list containing *once* each term present in the course overviews and apply the function `stem_hunspell` on this list. We save the terms and their stems in the tibble `stems`. Since the course descriptions can be conceptualized as a subset of the course overviews, the tibble `stems` contains all the terms present in the descriptions and can therefore safely used to stem the words in `d_description_tidy`.
```{r stemming dictionary, cache = TRUE}
stem_hunspell <- function(term) {
  stems   <- hunspell_stem(term)[[1]] # look up the term in the dictionary
  n_stems <- length(stems)            # number of stems
  if (n_stems == 0) term              # if no stems, return original term
  else              stems[[n_stems]]  # if multiple stems, return last (most basic) stem
}

dictionary <- d_overview_tidy %>% 
  select(word) %>%
  distinct %>%
  arrange(word) %>%
  mutate(word_stem = purrr::map_chr(.x = word,
                                    .f = stem_hunspell))

# terms for which the stem differs from the original word
filter(dictionary, word != word_stem)
```

```{r stemming d_description}
d_overview_stem <- d_overview_tidy %>%
  left_join(dictionary, by = "word") %>%
  rename(word_original = word,
         word = word_stem) %>%
  filter(!word %in% stop_words$word)
print(d_overview_stem) # See humanities - humanity

d_description_stem <- d_description_tidy %>%
  left_join(dictionary, by = "word") %>%
  rename(word_original = word,
         word = word_stem) %>%
  filter(!word %in% stop_words$word)
print(d_description_stem) # See starting - start
```
